\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

\newenvironment{proof}{\paragraph{Proof:}}{\hfill$\square$}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\author{Arthur Chen}
\title{Problem 12}
\date{\today}

\begin{document}

\section*{Problem 12}

Let $X_1 \dots X_n$ be iid from an exponential distribution with $f(x|\theta) = \theta e^{-\theta x}, x \geq 0$. Derive a likelihood test for $H_0: \theta = \theta_0$ versus $H_A: \theta \neq \theta_0$ and show that the rejection region is of the form $\bar{X} e^{-\theta_0 \bar{X}} \leq c$.

The likelihood ratio is

\[
\Lambda = \frac{\max_{\theta = \theta_0} lik(\theta)}
{\max_{\theta \in \mathbb{R^+}} lik(\theta)}
\]

where $lik(\theta|x) = \theta^n e^{-\theta \sum X_i}$. The maximum likelihood estimator of an exponential distribution is $\theta^* = \frac{1}{\bar{X}}$. Since small values of $\Lambda$ are evidence against the null, the test rejects the null when

\[
\Lambda = \frac{\theta_0^n e^{-\theta_0 n \bar{X}}}
{(\frac{1}{\bar{X}})^n e^n} < c
\]

$\theta_0^n$ is a constant and can be immediately absorbed into c. For the exponent, $\bar{X}^n e^{-\theta_0 n \bar{X} - n} = (\bar{X}e^{-\theta_0 \bar{X}-1})^n$, and since $\Lambda$ is positive, taking the nth root won't introduce more inequalities. Thus by taking the nth root of both sides, $e^{-1}$ can be absorbed into c, and so the rejection region has the form

\[
\bar{X} e^{-\theta_0 \bar{X}} < c
\]

as desired.

\section*{Problem 13}

Continuing from the previous problem, suppose that $\theta_0 = 1$, $n = 10$, and $\alpha = .05$.

\subsection*{Part a}

Show that the rejection region is of the form $\{\bar{X} \leq x_0\} \cup \{\bar{X} \geq x_1\}$, where $x_0, x_1$ are determined by $c$.

First note that $f(\bar{X}) = \bar{X}e^{-\theta_0 \bar{X}}$ is continuously differentiable. Note that $f(\bar{X})$ is always nonnegative, $f(0) = 0$, and $\lim_{\bar{X} \rightarrow \infty} f(\bar{X}) = 0$. Since $f'(\bar{X})$ is continuous, $f'(\bar{X}) = 0$ is a necessary condition for a local extrema. $f(\bar{X}) = 0$ when $\bar{X}^* = \frac{1}{\theta_0}$, and since this is the only inflection point of $f$, $f(\bar{X})$ must have a local extrema at $\bar{X}^*$. It's obvious that this is a maximum, since the endpoint limits go to zero.

Therefore, small values of $f(\bar{X})$ correspond to areas towards 0 and infinity, and away from the central hump. This corresponds to $\{\bar{X} \leq x_0\} \cup \{\bar{X} \geq x_1\}$.

\subsection*{Part b}

Explain why $c$ should be chosen so that $P(\bar{X} e^{-\bar{X}} \leq c) = .05$ when $\theta_0 = 1$.

Under the null hypothesis of $\theta_0 = 1$, the test statistic $\bar{X} e^{-\bar{X}}$ being small is evidence against the null. We want a 95 percent confidence set, so under the null, $P(reject null) = .05$. We reject the null when the test statistic is below a level c, so we need to choose c such that $P(reject null) = P(\bar{X} e^{-\bar{X}} \leq c) = .05$.

\subsection*{Part c}

Explain why $\sum_{i=1}^{10} X_i$ and $\bar{X}$ follow gamma distributions. Use this information to choose c.

It's well established that when $X$ and $Y$ are iid Gamma with the same rate parameters, their sum is a Gamma random variable with its rate parameter equal to X and Y's common rate parameter, and its shape parameter equal to the sum of X and Y's shape parameters. Since an exponential random variable is a $Gamma(1, \theta_0)$ random variable, $\sum_{i=1}^{10} X_i$ is distributed $Gamma(10, \theta_0)$.

Let $X$ be a Gamma random variable, and $Y = \frac{X}{m}$, with m a positive constant. Then $F_Y{y} = F_X{cy}$, and

\[
f_Y(y) = cf_X{cy} = m\frac{\theta^\alpha}{\Gamma(\alpha)} (my)^{\alpha-1} e^{-\theta(my)}
= \frac{(\theta m)^\alpha}{\Gamma(\alpha)} y^{\alpha-1} e^{-(\theta m)y}, y \geq 0
\]

which is a $Gamma(\alpha, \theta m)$ random variable. Thus, since $\sum_{i=1}^10 \sim Gamma(10, \theta_0)$, $\bar{X} = \frac{1}{10}\sum_{i=1}^10 \sim Gamma(10, 10\theta_0)$. Under the assumption that $\theta_0 = 1$, $\bar{X} \sim Gamma(10, 10)$.

Using this knowledge, we can use a transformation to find the distribution of $\bar{X}e^{-\bar{X}}$. However, I do not know how to do this, as the inverse of $\bar{X}e^{-\bar{X}}$ is $-W(-\bar{X})$ according to Wolfram Alpha, where $W(x)$ is the Lambert W function. I do not know enough about this function at the moment to numerically determine c.

\subsection*{Part d}

Explain how to approximate c by simulation.

Since we know the distribution of $\bar{X}$, we can approximate the distribution of $\bar{X}e^{-\bar{X}}$. First we generate n $Gamma(10, 10)$ variables, apply $f(x) = xe^{-x}$ to the generated numbers, and rank them from smallest to largest. These numbers estimate the distribution of $\bar{X}e^{-\bar{X}}$, and so the $.05n$th number from the bottom is our estimate for c such that $P(\bar{X}e^{-\bar{X}} < c)$.

\end{document}