---
title: "Estimating the Paramater of the Poisson Distribution with MLE and Bayes"
output: pdf_document
---

```{r setup, include = FALSE}
library(reticulate)

```

```{r, echo=FALSE}
data <- read.table("../Excel Comma/Chapter 8/poisson-asbestos.txt", quote="\"", comment.char="")
data <- data$V1
```

## Maximum Likelihood Estimation

For a Poisson distribution, the MLE estimator matches the method of moments estimator with $\hat{\lambda} = \bar{X}$, while the standard error of $\hat{\lambda}$ is $s_{\hat{\lambda}} = \sqrt{\frac{\hat{\lambda}} {n}}$.

```{r}
mleLambdaBar <- round(mean(data), 1)
mleSLambdaBar <- round(sqrt(mleLambdaBar / length(data)), 2)
```

Thus the MLE estimates of $\bar{\lambda}$ and $s_{\bar{\lambda}}$ are `r mle.lambdaBar` and `r mle.sLambdaBar`, respectively, as in the book.

## Bayesian Estimation, Uniform Prior

Here, we assume that the prior distribution of $\lambda$ follows a uniform distribution on [0, 100]. Thus, the posterior density of $\lambda$ takes the form

$$ 
f_{\Lambda | X}(\lambda | x) \propto \lambda^{\sum^{n}_{i=1} x_i} e^{-n\lambda}, 0 \leq \lambda \leq 100
$$

First we compute the numerator of the posterior density, up to a proportionality constant. To prevent the production of infinities, we will shrink the posterior density of lambda by a factor of $10^{-500}$.

```{r}
bayesPostDensNum <- function(lambda) {
  xsum <- sum(data)
  n <- length(data)
  return(
    exp(
      -500 * log(10) + 
        (xsum * log(lambda) - (n * lambda))
    )
  )
}
```

```{r}
plot.function(bayesPostDensNum, from = 0, to = 100, n = 400)
```
