---
title: "Problem 43"
output: pdf_document
---

# Gamma Ray Interarrival Times

## Initial Histogram

```{r, echo=FALSE}
data <- read.table("../Excel Comma/Chapter 8/gamma-arrivals.csv", quote="\"", comment.char="")
data <- data[[1]]
```

The data consists of `r length(data)` observations of interarrival times for gamma rays, with units in seconds.

```{r}
hist(data, breaks = seq(from = 0, to = 750, by = 1), main = 'Histogram of Interarrival Times', xlab = 'Time (seconds)')
```
The interarrival times look roughly exponential, or at least gamma, since the exponential distribution is a special case of the gamma distribution. For the moment, we will assume that interarrival times are distributed via a gamma distribution.

## Parameter Estimates

### Method of Moments

The method of moments estimators for $\alpha$ and $\gamma$ are

$$ 
\hat{\alpha}_{MoM} = \frac{\bar{X}^2}{\bar{\sigma}^2}
$$
$$
\hat{\lambda}_{MoM} = \frac{\bar{X}}{\bar{\sigma}^2}
$$

```{r}
aMoM <- (mean(data)^2) / var(data)
lMoM <- mean(data) / var(data)
```

Thus the method of moments estimates are $\hat{\alpha}_{MoM}$ = `r aMoM` and $\hat{\lambda}_{MoM}$ = `r lMoM`.

## Maximum Likelihood

The maximum likelihood estimator for $\alpha$ satisfies the following equation

$$
nln(\hat{\alpha}) - nln(\bar{X}) + \sum_{i=1}^n ln(X_i) - n \psi(\hat{\alpha}) = 0
$$

Where $\psi(x)$ is the digamma function, $\frac{\Gamma'(x)}{\Gamma(x)}$.
