---
title: "Problem 43"
output: pdf_document
---

# Gamma Ray Interarrival Times

## Initial Histogram

```{r, global_options, include=FALSE}
knitr::opts_chunk$set(tidy = TRUE, tidy.opts = list(width.cutoff=55))
```

```{r, echo=FALSE}
data <- read.table("../Excel Comma/Chapter 8/gamma-arrivals.csv", quote="\"", comment.char="")
data <- data[[1]]
```

The data consists of `r length(data)` observations of interarrival times for gamma rays, with units in seconds.

```{r}
hist(data, breaks = seq(from = 0, to = 750, by = 1), main = 'Interarrival Times', xlab = 'Time (seconds)')
```
The interarrival times look roughly exponential, or at least gamma, since the exponential distribution is a special case of the gamma distribution. For the moment, we will assume that interarrival times are distributed via a gamma distribution.

## Parameter Estimates

### Method of Moments

The method of moments estimators for $\alpha$ and $\lambda$ are

$$ 
\hat{\alpha}_{MoM} = \frac{\bar{X}^2}{\bar{\sigma}^2}
$$
$$
\hat{\lambda}_{MoM} = \frac{\bar{X}}{\bar{\sigma}^2}
$$

```{r}
#Computes the MoM estimates of alpha and lambda for a gamma distribution, respectively
gammaMoM <- function(data) {
  output <- list()
  output$alpha <- (mean(data)^2) / var(data)
  output$lambda <- mean(data) / var(data)
  return(output)
}
foo <- gammaMoM(data)
aMoM <- foo$alpha
lMoM <- foo$lambda
rm(foo)
```

Thus the method of moments estimates are $\hat{\alpha}_{MoM}$ = `r aMoM` and $\hat{\lambda}_{MoM}$ = `r lMoM`.

### Maximum Likelihood

The maximum likelihood estimator for $\alpha$ satisfies the following equation

$$
nln(\hat{\alpha}) - nln(\bar{X}) + \sum_{i=1}^n ln(X_i) - n \psi(\hat{\alpha}) = 0
$$

Where $\psi(x)$ is the digamma function, $\frac{\Gamma'(x)}{\Gamma(x)}$. Then an estimator for $\lambda$ is

$$
\hat{\lambda}_{MLE} = \frac{\hat{a}}{\bar{X}}
$$

We will use the method of moments estimator as an initial guess to solve for $\hat{\alpha}_{MLE}$.

```{r}
#Computes the MLE estimates of alpha and lambda for a gamma distribution, respectively
gammaMLE <- function(data) {
  aMoM <- gammaMoM(data)$alpha
  #Initialize variables
  n <- length(data)
  lnXBar <- log(mean(data))
  sumLnX <- sum(log(data))
  aMLEFun <- function(a) {
    n*log(a) - n*lnXBar + sumLnX - n*digamma(a)
  }
  
  #diagnostic plots
  #x <- seq(from = aMoM - .5, to = aMoM + .5, by = 0.01)
  #plot(x, aMLEFun(x))
  output = list()
  output$alpha <- uniroot(aMLEFun, interval = c(aMoM - .5, aMoM + .5))$root
  output$lambda <- output$alpha / mean(data)
  return(output)
}

foo <- gammaMLE(data)
aMLE <- foo$alpha
lMLE <- foo$lambda
rm(foo)
```

Thus the maximum likelihood estimates are $\hat{\alpha}_{MLE}$ = `r aMLE` and $\hat{\lambda}_{MLE}$ = `r lMLE`. We see that the two are nearly identical. Looking at the estimated distributions versus the histogram shows that the fit is good, and that the two estimated densities are nearly identical.

```{r}
hist(data, freq = FALSE, breaks = seq(from = 0, to = 750, by = 1), density = 1, main = 'Probability Densities of Interarrival Times', xlab = 'Time (seconds)')
xSeq <- seq(from = 0, to = 750, by = 1)
#MoM fit
lines(xSeq, dgamma(xSeq, shape = aMoM, rate = lMoM), col = 'blue', lwd = 1)
#MLE fit
lines(xSeq, dgamma(xSeq, shape = aMLE, rate = lMLE), col = 'red', lwd = 1)
legend('topright', legend = c('MoM', 'MLE'), col=c('blue', 'red'),
       lty=1, cex=0.8)
```

## Estimating Standard Errors via Bootstrap

```{r}
#First step in a bootstrap. Generates simulated gamma random variables.
#Creates dataSize iid draws from a gamma distribution, and estimate alpha and lambda hat using calcFun. Repeat this rep times.
#Returns a list with alphaHats, lambdaHats being the respective estimates
gammaSim <- function(dataSize, reps, alpha, lambda, calcFun) {
  output <- list()
  output$alphaHats <- vector(mode = 'numeric', length = reps)
  output$lambdaHats <- vector(mode = 'numeric', length = reps)
  
  for (i in 1:reps) {
    data <- rgamma(dataSize, shape = alpha, rate = lambda)
    ests <- calcFun(data)
    output$alphaHats[i] <- ests$alpha
    output$lambdaHats[i] <- ests$lambda
  }
  
  return(output)
}


#Bootstrap estimation of standard errors of alpha and lambda of a gamma distribution calculated uses calcFun (MoM or MLE).
gammaSE <- function(dataSize, reps, alpha, lambda, calcFun) {
  ests <- gammaSim(dataSize, reps, alpha, lambda, calcFun)
  
  #Return the estimated standard errors
  output <- list()
  output$alphaSE <- sd(ests$alphaHats)
  output$lambdaSE <- sd(ests$lambdaHats)
  return(output)
}
```

```{r}
set.seed(1000)
MoMSE <- gammaSE(dataSize = length(data), reps = 1000, alpha = aMoM, lambda = lMoM, calcFun = gammaMoM)
MLESE <- gammaSE(dataSize = length(data), reps = 1000, alpha = aMLE, lambda = lMLE, calcFun = gammaMLE)
```

The estimates for $s_{\hat{\alpha}_{MoM}}$ and $s_{\hat{\lambda}_{MoM}}$ are thus `r MoMSE$alpha` and `r MoMSE$lambda`. The corresponding estimates for $s_{\hat{\alpha}_{MLE}}$ and $s_{\hat{\lambda}_{MLE}}$ are thus `r MLESE$alpha` and `r MLESE$lambda`. The estimated standard errors for the maximum likelihood estimates are significantly lower than the corresponding estimates for the method of moments.

## Estimating Confidence Intervals via Bootstrap

```{r}
#Computes the bootstrapped 100(1-confidenceAlpha) confidence intervals of gamma distribution parameters calculated via calcFun.
#Note the confidenceAlpha should be a decimal, not a percent.
#Returns a list with alphaCI and lambdaCI being the confidence intervals.
gammaCI <- function(dataSize, reps, alpha, lambda, calcFun, confidenceAlpha) {
  #Convert confidenceAlpha to a decimal if it was accidentally inputted as a percent
  if (confidenceAlpha > 1) {
    confidenceAlpha = confidenceAlpha / 100
  }
  ests <- gammaSim(dataSize, reps, alpha, lambda, calcFun)
  
  #Order the lists
  sortedAlpha <- sort(ests$alphaHats)
  sortedLambda <- sort(ests$lambdaHats)
  
  low <- floor((confidenceAlpha / 2) * reps)
  high <- reps - low
  
  output <- list()
  output$alphaCI <- c(sortedAlpha[low], sortedAlpha[high])
  output$lambdaCI <- c(sortedLambda[low], sortedLambda[high])
  return(output)
}
```

```{r}
set.seed(2000)
MoMCI <- gammaCI(dataSize = length(data), reps = 1000, alpha = aMoM, lambda = lMoM, calcFun = gammaMoM, confidenceAlpha = 0.05)
MLECI <- gammaCI(dataSize = length(data), reps = 1000, alpha = aMLE, lambda = lMLE, calcFun = gammaMLE, confidenceAlpha = 0.05)
```

The estimated 95% confidence intervals for $\hat{\alpha}_{MoM}$ and $\hat{\lambda}_{MoM}$ are (`r MoMCI$alphaCI[1]`, `r MoMCI$alphaCI[2]`) and (`r MoMCI$lambdaCI[1]`, `r MoMCI$lambdaCI[2]`), respectively. The corresponding estimates for $\hat{\alpha}_{MLE}$ and $\hat{\lambda}_{MLE}$ are (`r MLECI$alphaCI[1]`, `r MLECI$alphaCI[2]`) and (`r MLECI$lambdaCI[1]`, `r MLECI$lambdaCI[2]`), respectively. As we can see, the parameter estimates estimated via method of moments are much less dispersed than the corresponding estimates from method of moments.

The interarrival times are consistent with a Poisson arrival process. In a Poisson arrival process, interarrival times are exponentially distributed, which corresponds to a gamma distribution with $\alpha = 1$. As we can see by the 95% confidence intervals, we can not reject the null hypothesis that $\alpha = 1$ for neither the method of momentsn nor maximum likelihood estimates.
