---
title: "Problem 42"
output: pdf_document
---

## Estimating the Parameter of Poisson Arrivals with Variable Time Lengths

```{r}
data <- read.csv("../Excel Comma/Chapter 8/gamma-ray.csv")
```

We will assume that gamma ray arrivals follow a Poisson process with parameter $\lambda$ arrivals per second. Under this assumption, let the `r length(data[[1]])` observations be indexed as $i = 1, 2, ... n$. Let $t_i$ and $X_i$ be the length of the $i$th time interval and the number of observed gamma rays during the $i$th time interval, respectively. I will assume that the time intervals are disjoint.

According to the model, $X_i$ follows a Poisson distribution with parameter $\lambda t_i$. Due to the assumption that the time intervals are disjoint, the $X_i$s are independent of each other.

The distribution of $X_i$ is thus

$$
P(X_i = k_i | \lambda, t_i) = \frac{(\lambda t_i)^{k_i} e^{- \lambda t_i}}
  {k_i!}
$$

The likelihood function of $\lambda$ is thus

$$
lik(\lambda|\textbf{t, X }) = \frac{
    (\lambda t_i)^{\sum_{i=1}^{n} X_i} e^{- \lambda \sum_{i=1}^{n} t_i}
  }
  {\prod_{i=1}^{n} (X_i!)}
$$

The log likelihood is thus

$$
l(\lambda) = ln(\lambda) \sum_{i=1}^n X_i + \sum_{i=1}^n X_i ln(t_i)
  - \lambda \sum_{i=1}^n t_i - \sum_{i=1}^n ln(X_i !)
$$

The maximum likelihood estimator satisfies the first-order condition

$$
0 = \frac{1}{\hat{\lambda}} \sum_{i=1}^n X_i - \sum_{i=1}^n t_i
$$

Which after some rearranging, yields the maximum likelihood estimator

$$
\hat{\lambda} = \frac{\sum_{i=1}^n X_i}{\sum_{i=1}^n t_i}
$$
Setting all $t_i$'s to 1 yields the usual estimator for $\lambda$ in an iid sample from a Poisson distribution.

```{r}
lambdaHat <- sum(data$count) / sum(data$seconds)
```

Thus $\hat{\lambda}$ = `r lambdaHat`.

## The Standard Error of the Estimate

Define $S = \sum_{i=1}^n X_i$ and $T = \sum_{i=1}^n t_i$. Then the estimator $\hat{\lambda}$ can be rewritten as 

$$
\hat{\lambda} = \frac{S}{T}
$$

Since $S$ is the sum of independent Poisson variables with parameter $\lambda t_i$, $S$ follows a Poisson distribution with parameter $\lambda T$. Thus $Var(S) = \lambda T$. Since $\hat{\lambda} = \frac{S}{T}$, $Var(\hat{\lambda}) = \frac{\lambda}{T}$, and $SD(\hat{\lambda}) = \sqrt{\frac{\lambda}{T}}$. This gives us an estimator

$$
s_{\hat{\lambda}} = \sqrt {\frac{\hat{\lambda}}{T}}
$$
```{r}
s_lambdaHat <- sqrt(lambdaHat / sum(data$seconds))
```

This produces $s_{\hat{\lambda}}$ = `r s_lambdaHat`.

